DRAFT: Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network

### 2017-02-28 preprint/source/paper.tex
Add a bunch of stuff.

%\usepackage{draftwatermark}
%\SetWatermarkText{DRAFT}

@@

% Copyright
\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\doi{10.475/123_4}

% ISBN
%\isbn{123-4567-24-567/08/06}

%Conference
\conferenceinfo{WWW '17}{April 3--7, 2017, Perth, Australia}

\acmPrice{\$15.00}

@@

\title{Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
Networks with single points of failure are particularly susceptible to
targeted attacks.
In communication networks, these types of faults can leave users vulnerable to
censorship and targeted surveillance,
even when cryptography is utilized.
Centralized networks have single points of failure by definition,
leading to a growing popularity in decentralized architectures and protocols.
However, centralized network structure can arise even when protocols are
decentralized.
While based on decentralized protocols, the Internet and World-Wide Web have
been shown both theoretically and historically to be highly susceptible to
adversarial faults, in part due to emergent, structural centralization.
Existing network trust models, such as webs of trust, fail to adequately address
network structure.
We describe a novel, adversarial fault-tolerant,
concurrent multipath routing algorithm for the decentralized
butterfly network topology.
We also develop a partial trust model that makes it possible to
quantify the adversarial fault tolerance of a network,
while also making more realistic transitivity assumptions than webs of trust.
When network topology can be dictated, these results can be used to create
scalable, attack-tolerant infrastructures. More generally, our results provide
a formalism for evaluating the effects of network structure on adversarial
fault tolerance.
\end{abstract}

@@

@@

Large-scale communication networks, exemplified by the Internet,
have become ubiquitous, and are now crucial infrastructure for
individuals, communities, and organizations around the world.
As with any critical infrastructure, the cost of a failure can be
immense, so methods for tolerating various kind of faults are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
In fact, the decentralized design of the Internet was motivated
by the need to withstand nuclear strikes
\cite{baran_distributed_1964}.
{\em Adversarial faults},
in which an adversary can strategically target attacks,
are some of the most difficult to protect against.
Such attacks are often used for censorship and targeted surveilance.
Both theoretical results and recent events (described below)
have demonstrated that
the Internet is surprisingly vulnerable to adversarial faults.
Decentralization remains a promising approach to building resilient networks,
but further work is needed to understand the relationship between
decentralized network structure and adversarial fault tolerance.

@@

The Internet's vulnerability to censorship and other targeted attacks
has been confirmed by the success of several such attacks.
For example, in 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
Such vulnerabilities are not limited to any one system or protocol,
but result from the network structure itself.
With strong theoretical and historical evidence that network structure
can create vulnerabilities,
methods for analyzing structural vulnerabilities and for designing
fault tolerant networks are needed.
This paper presents several contributions towards advancing those goals.

We consider a setting in which a source node (Alice)
in a network attempts to route a message
to a target node (Bob) by forwarding it through the links of the network.
We assume that some nodes in the network may be compromised by an attacker
(Mal).
Compromised nodes may behave incorrectly by blocking, altering,
or incorrectly routing messages.
We assume that Mal has full knowledge of the network structure, but has
limited resources and thus can only compromise a fixed number of nodes.
We also assume that nodes within a fixed network
distance of either Alice or Bob are
{\em trusted}
and cannot be compromised.
This assumption,
which we call
{\em partial trust transitivity},
is a weaker (more realistic) form of the transitive
trust assumption used by the popular 
web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}.

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance,
and propose a method for achieving a high-level of attack-tolerance on
the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure and high connectivity.
While it may seem questionable to dictate the topology of trust relationships,
there are many applications where this occurs,
e.g., overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
Furthermore, we stress that trust is not static, it changes over time
(ableit slowly).
In this sense, we approach trust as deliberate infrastructure.
Since centralized network structure is inherently
vulnerable to attack, the ability to deliberately structure trust relationships
is a necessary component of any attack-tolerance scheme.

We first describe how fault tolerance techniques can be adapted and
evaluated in a network setting with partial trust.
Generally, faults in network paths can be correlated,
preventing the appliation of standard fault tolerance techniques
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
which assume independent faults.
By using {\em independent paths},
which have no untrusted nodes in common,
we model communication across a complex network by a number of
redundant and independent channels.
Fault-tolerance schemes can then be applied when redundant messages
are sent in parallel,
a technique known as {\em concurrent multipath routing}
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013}.
The receiver can then use the redundant messages to detect and/or correct
errors.
We formally evaluate the
effects of network structure on attack-resistance and show that the probability
of undetected errors decreases exponentially with the number of
independent paths between source and destination,
even when no individual path is entirely trusted.

We also propose a novel concurrent multipath routing algorithm for the butterfly
topology.
The algorithm identifies independent paths,
which when combined with the fault-tolerant concurrent multipath routing
scheme above,
achieves a high level of adversarial fault tolerance on the butterfly topology.

@@

This paper is organized as follows.
Section \ref{sec-related}. reviews background and related work.
Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
infrastructure.
Section \ref{sec-ft}. describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly}. describes a routing algorithm for multipath fault tolerance
on the butterfly network topology.
Section \ref{sec-discussion}. discusses the our results.
And Section \ref{sec-conclusion}. concludes.

@@

There has been considerable work on trust in network security.
Both centralized and decentralized approaches are commonly used to create
trust infrastructures.
Centralized approaches such as {\em public key infrastructure} (PKI)
suffer from a number of vulnerabilities
\cite{ellison_ten_2000},
which stem largely from the single points of failure inherent to
centralization.
The well-known and widely-used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003}
is a decentralized alternative.
In a web of trust,
individuals choose who they trust initially.
Trust is then extended to new individuals if they are vouched for by a
currently-trusted individual,
making it possible to extend their web of trust to a large
number of nodes.
However, the web of trust approach assumes that trust can be extended
transitively,
which is unrealistic
\cite{christianson_why_1997}.
Previous work on evaluating network structure has focused on authentication
protocols, showing that independent paths can reduce an adversary's ability
to impersonate a target
\cite{levien_attack-resistant_2009}.
Other work has shown that identifying independent paths in arbitrary networks
is NP-hard and provided approximation algorithms
\cite{reiter_resilient_1998}.
Our work complements these by explicitly quantifying the effects of partial
trust transitivity,
and providing a routing algorithm to efficiently construct independent
paths on a structured network.

Many distributed consensus protocols (such as those used by cryptocurrencies)
are designed to provide tolerance against arbitrary or adversarial faults.
Byzantine agreement protocols
\cite{lamport_byzantine_1982,castro_practical_1999}
provide tolerance against arbitrary faults (including attacks) under
some circumstances, but are limited to small networks due to poor scalability.
Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
and proof-of-stake \cite{king_ppcoin:_2012}
provide better scalability,
but are wasteful of computational and energy resources.
Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
is scalable, allows for flexible trust,
and is highly fault-tolerant on networks meeting a set of requirements.
However, FBA does not provide a method for evaluating the
fault tolerance properties of different network structures
or calculating how the failure probability for a particular network.

Most existing adversarial fault tolerance schemes focus on end-to-end,
protocol-based, cryptographic solutions
\cite{ferguson_practical_2003},
without addressing single points of failure in-between
(i.e., {\em man-in-the-middle attacks}).
There are however, a few notable fault tolerance schemes
that focus on network structure.
Fiat and Saia described a scheme that combines the butterfly topology
with expander graphs to create a highly censorship-resistant content-addressable
network \cite{fiat_censorship_2002},
although this scheme exhibits poor scaling due to a very high level of data
replication.
Perhaps the most mature structural solution is the Freenet collaboration
\cite{clarke_freenet:_2001}.
Freenet uses secret sharing
\cite{shamir_how_1979, blakley_safeguarding_1979}
and small-world routing
\cite{zhang_using_2002,kleinberg_small-world_2000}
to create a content-addressable network with a high level of both
confidentiality and censorship resistance.
Freenet guarantees that data is stored redundantly,
but still allows for centralized network structure and thus single points
of failure as data travels from its origin to the redundant storage locations.

@@

Our proposed routing algorithm makes use of a
{\em structured network}, in which link structure is predetermined.
Such networks can be designed to have favorable structural and routing
properties,
at the expense of complicating the addition or removal of nodes.
Structured networks have been a popular tool in parallel processing
architectures \cite{kshemkalyani_distributed_2008}.
More recently, peer-to-peer systems based on distributed hash tables have used
structured ``overlay'' networks to map table keys to local TCP/IP routes
\cite{lua_survey_2005,korzun_structured_2013}.

@@

\begin{description}
\item[Scalability:]
For large scale networks it is important that the infrastructure allows
for the network to grow while remaining functional.
In practice, people, devices, and connections, have limited capabilities
and these limitations need to be considered as part of the design of the
infrastructure.

\item [Decentralization:]
Systems having single points of failure are less tolerant against faults at
those points.
The existence of such points not only increases the likelihood that an attack
will succeed,
but also incentivizes attack by presenting effective targets.
In order to minimize single points of failure,
attack tolerant infrastructures must use decentralized protocols and
decentralized network structures.

\item[Stabilizing asymmetry:]
In the context of international conflict,
{\em asymmetric conflicts} are a special case that makes it possible for the
less powerful party to have an advantage over the more powerful party
\cite{mack_why_1975}.
In asymmetric conflicts, the same level of resource expenditure yields different
results for different parties;
the attacker's resources are either more or less effective than the defender's.
We call the latter case stabilizing asymmetry,
because it reduces an attacker's power relative to their target.
With this in mind, an attack-resistant infrastructure will benefit from a high
level of stabilizing asymmetry.

@@

We are specifically concerned with network structure-based approaches to
fault tolerance.
In networks, specific structural properties are required to achieve the
functional properties described above.
\begin{description}
\item [Sparsity and low diameter:]
To achieve scalability, networks must be {\em sparse} and have a
{\em low diameter}.
In practical settings, humans and devices have an upper limit on the number
of connections they can maintain (e.g., Dunbar's number
\cite{dunbar_neocortex_1992}).
In sparse networks, the number of links grows slowly as the network grows in
size, allowing the network to scale without exceeding the nodes' capacity for
links.
Similarly, low-diameter guarantees that as a network grows, a short path will
still exist between any pair of nodes.
While low diameter guarantees a path exists,
paths are only useful if an efficient {\em routing} algorithm exists
to find them.

@@

\item [Redundancy:]
In a network, redundancy refers to the existence of multiple non-overlapping
paths between nodes or components.
Redundancy can help reduce single points of failure and decrease
centralization.
One measure of redundancy is given by the ratio of edges to nodes
\cite{baran_distributed_1964}.
A single point of failure occurs when a node holds a uniquely
central position.
When alternative paths are added to bypass central nodes,
the network becomes more redundant.

@@

Within the field of {\em fault tolerance},
many techniques have been developed for building reliable systems
out of unreliable components
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
We will make use of standard fault tolerance terminology, summarized here.
A {\em fault} occurs when one component
of a system behaves incorrectly (e.g., a routing node blocking or
altering a message).
The result of that fault (e.g., a recipient receiving conflicting messages)
is an {\em error} state.
If the error is undetected or corrected to the wrong value, the system is
said to have experienced a {\em failure} (e.g., an altered message is
accepted as authentic).
We are concerned in particular with {\em adversarial faults},
which (as opposed to random faults)
are chosen strategically to maximize the likelihood of a failure.

@@

An important challenge in large-scale, secure communication is the ability to
communicate reliably and securely with remote parties for whom no direct,
trusted link exists.
The commonly-used web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}.
extends trust infinitely transitively,
to friends of friends, and friends of friends of friends, and so on.
However, the assumption of infinitely transitive trust is unrealistic
\cite{christianson_why_1997},
and does not allow for the analysis of network structure.

We can allow for more realistic transitivity assumptions and incorporate
network structure into a trust model with one key insight:
even if no single path between a sender and receiver is fully trusted,
multiple copies of a message can be sent along different paths and compared
by the receiver to detect and correct errors.
However, in the presence of an attacker, there is only benefit in sending an
additional message along a path if that path does not share untrusted
single points of failure with any of the existing message paths
(otherwise, the adversary can compromise both messages
by causing a single fault).
Following \cite{reiter_resilient_1998},
we call paths that do not share any untrusted single points of failure
{\em independent}.
The maximum number of independent paths represents the effective redundancy that
can be utilized by any redundancy-based fault tolerance scheme.
We now propose a method for quantifying the effective redundancy of a network
as a function of its structure and level of trust transitivity.
For the purpose of analyzing adversarial fault tolerance,
this method allows a communication network
to be modeled as a set of virtual links that directly
connect each node pair,
with each virtual link providing some level of redundancy.

We now specify the {\em partial trust model}.
We assume an undirected graph $G = (V,E)$,
although the model can easily be extended to directed multigraphs.
Vertices representing commiunicating agents,
and with edges representing mutually trusted communication links.
Let $v \in V$ be an arbitrary sender (Alice)
and $w \in V$ be an arbitrary receiver (Bob).
We assume the presence of an adversary (Mal) who knows the
full structure of the network,
and who can compromise a fixed number of agents,
gaining complete control of their behavior,
as long as those agents are not trusted by either Alice or Bob.
We define a {\em trust radius} $h$ and that nodes $u$ and
$u^\prime$ trust each other if their distance is less than $h$.
For a given node $u$,
we call the set of trusted nodes its
{\em trusted neighborhood} $T_h(u)$,
and all nodes at exactly distance $h$ the
{\em trust boundary} $B_h(u)$:
\beq
T_h(u) &=& \left\{ u \mid d(u,u^\prime) < h \right\} \\
B_h(u) &=& \left\{ u \mid d(u,u^\prime) = h \right\}.
\eeq
The trust boundary $B_h$ plays an important role because these nodes are not
trusted by $u$,
but if compromised can entirely isolate $u$ from the rest of the network.
These trust assumptions imply that when Alice sends a message to Bob,
Mal can only cause faults in the set of nodes outside both of their trusted
neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
We refer to this set of nodes as the {\em untrusted region}.

Having defined the assumptions of the partial trust model,
we now quantify the effective redundancy between Alice and Bob
when trust radius $h$ is assumed.
This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
the graph after Alice's and Bob's trusted neighborhoods have been
collapsed into single source/sink vertices.
Each trust boundary forms a cut of the network and places an upper bound on the
min-cut:
\beq
\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
\eeq
Equality holds when there are no bottlenecks within the untrusted region,
an indication that the network is decentralized.
The redundancy of the entire graph can be characterized by the minimum over
all vertex pairs:
\beq
\delta_h(V) \equiv \min_{v,w \in V} \delta_{v,w,h}.
\eeq
Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
redundant paths can be constructed between them.
While $\delta_h$ is a purely structural property of the graph and
may have additional applications,
we are interested in this quantity because
it places an upper bound on the effectiveness of any
redundancy-based fault tolerance scheme.
As a function of $h$, $\delta_h$ describes a network's ability to create
redundancy from trust.
Combined with a {\em routing algorithm} that tells nodes how to forward
messages to guarantee they take independent paths,
this redundancy can be used to achieve fault tolerance.
For attack-tolerant networks,
it is thus desirable for $\delta_h$ to increase quickly as a function of $h$.
Furthermore, by applying the partial trust model,
an adversarial fault tolerance scheme can treat each pair of nodes as being
directly connected by a virtual link with least $\delta_h$ redundant
channels.
We will now describe such a scheme and evaluate its fault tolerance properties.

>>>>

@@

% Copyright
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
\doi{}

% ISBN
\isbn{}

%Conference
\conferenceinfo{Pre-print}{Currently under peer review.}

@@

\title{DRAFT: Towards Attack-Tolerant Networks: Concurrent Multipath Routing and the Butterfly Network}
%
% You need the command \numberofauthors to handle the 'placement
% and alignment' of the authors beneath the title.
%
% For aesthetic reasons, we recommend 'three authors at a time'
% i.e. three 'name/affiliation blocks' be placed beneath the title.
%
% NOTE: You are NOT restricted in how many 'rows' of
% "name/affiliations" may appear. We just ask that you restrict
% the number of 'columns' to three.
%
% Because of the available 'opening page real-estate'
% we ask you to refrain from putting more than six authors
% (two rows with three columns) beneath the article title.
% More than six makes the first-page appear very cluttered indeed.
%
% Use the \alignauthor commands to handle the names
% and affiliations for an 'aesthetic maximum' of six authors.
% Add names, affiliations, addresses for
% the seventh etc. author(s) as the argument for the
% \additionalauthors command.
% These 'additional authors' will be output/set for you
% without further effort on your part as the last section in
% the body of your article BEFORE References or any Appendices.

@@

\maketitle
\begin{abstract}
Networks with single points of failure are particularly susceptible to
targeted attacks.
In communication networks, these types of faults can leave users vulnerable to
censorship and targeted surveillance,
even when cryptography is utilized.
Centralized networks have single points of failure by definition,
leading to a growing popularity in decentralized architectures and protocols.
However, centralized network structure can arise even when protocols are
decentralized.
Despite being based on decentralized protocols,
the Internet and World-Wide Web have
been shown both theoretically and historically to be highly susceptible to
adversarial faults, in part due to emergent structural centralization.
Existing network trust models, such as webs of trust, fail to adequately address
network structure.
We describe a novel, adversarial fault-tolerant,
concurrent multipath routing algorithm for the decentralized
butterfly network topology.
We also develop a partial trust model that makes it possible to
quantify the adversarial fault tolerance of a network,
and which makes more realistic transitivity assumptions than webs of trust.
When network topology can be dictated, these results can be used to create
scalable, attack-tolerant infrastructures. More generally, our results provide
a formalism for evaluating the effects of network structure on adversarial
fault tolerance.
\end{abstract}

@@

\vskip 3em

@@

Large-scale communication networks, exemplified by the Internet,
have become ubiquitous, and are now crucial infrastructure for
individuals, communities, and organizations around the world.
As with any critical infrastructure, the cost of a failure can be
immense, so methods for tolerating various kind of faults are an
important and ongoing area of research
\cite{zin_survey_2015,albert_error_2000,sterbenz_resilience_2010}.
In fact, the decentralized design of the Internet was motivated
by the need to withstand nuclear strikes
\cite{baran_distributed_1964}.
{\em Adversarial faults},
in which an adversary can strategically target attacks,
are some of the most difficult to protect against.
Such attacks are often used for censorship and targeted surveilance.
Both theoretical results and recent events (described below)
have demonstrated that
the Internet is surprisingly vulnerable to adversarial faults.
Decentralization remains a promising approach towards
building resilient networks,
but further work is needed to understand the relationship between
decentralized network structure and adversarial fault tolerance.

The Internet's vulnerability to censorship and other targeted attacks
has been confirmed by the success of several such attacks.
For example, in 2008, YouTube suffered a worldwide outage for several hours
when a service provider in Pakistan advertised false routing information
\cite{hunter_pakistan_2008}.
The action (known as a {\em black hole attack}) was intended to censor YouTube
within Pakistan only, but resulted in a worldwide cascading failure.
Such vulnerabilities are not limited to any one system or protocol,
but result from the network structure itself.

@@

With strong theoretical and historical evidence that network structure
can create vulnerabilities,
methods for analyzing structural vulnerabilities and for designing
fault-tolerant networks are needed.
This paper presents several contributions towards advancing those goals.

We consider a setting in which a source node (Alice)
in a network attempts to route a message
to a target node (Bob) by forwarding it through the links of the network.
A ``link'' in this context could represent any kind of connection
(e.g., physical cables, encrypted channels).
We assume that some nodes in the network may be compromised by an attacker
(Mal).
Compromised nodes may behave incorrectly by blocking, altering,
or incorrectly routing messages.
We assume that Mal has full knowledge of the network structure, but has
limited resources and thus can only compromise a fixed number of nodes.
We also assume that nodes within a fixed network
distance of either Alice or Bob are
{\em trusted}
and cannot be compromised.
This assumption,
which we call
{\em partial trust transitivity},
is a weaker (more realistic) form of the transitive
trust assumption used by the popular 
web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}.

Under the above assumptions,
we show how to evaluate the influence of network structure
on attack-tolerance,
and propose a method for achieving a high-level of attack-tolerance on
the butterfly network topology.
The butterfly topology is popular in parallel processing
\cite{kshemkalyani_distributed_2008} and
peer-to-peer \cite{lua_survey_2005, korzun_structured_2013}
applications, due to its regular structure and high connectivity.
While it may seem questionable to dictate the topology of trust relationships,
there are many applications where this occurs,
e.g., overlay networks
\cite{lua_survey_2005, korzun_structured_2013},
formal organizations \cite{mohr_explaining_1982},
government-regulated cellular networks \cite{walker_mass_2012},
and call tree notification systems \cite{nickerson_thinking_2010}.
Furthermore, we stress that trust is not static;
it changes over time (ableit slowly).
In this sense, we approach trust as deliberate infrastructure.
Since centralized network structure is inherently
vulnerable to attack, the ability to deliberately structure trust relationships
is a necessary component of any attack-tolerance scheme.

We first describe how fault tolerance techniques can be adapted and
evaluated in a network setting with partial trust.
Generally, faults in network paths can be correlated,
preventing the appliation of standard fault tolerance techniques
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956},
which assume independent faults.
By using {\em independent paths},
which have no untrusted nodes in common,
we model communication across a complex network by a number of
redundant and independent channels.
Redundant messages can be sent across these channels in parallel,
a technique known as {\em concurrent multipath routing}
\cite{zin_survey_2015, qadir_exploiting_2015, khiani_comparative_2013},
and used for fault tolerance.
The receiver can then use the redundant messages to detect and/or correct
errors.
We formally evaluate the
effects of network structure on attack-tolerance and show that the probability
of an undetected error decreases exponentially with the number of
independent paths between source and destination,
even when no individual path is entirely trusted.

We also propose a novel concurrent multipath routing algorithm for the butterfly
topology.
The algorithm identifies independent paths,
which can combined with the fault-tolerant concurrent multipath routing
scheme above to
achieve a high level of adversarial fault tolerance on the butterfly topology.

@@

This paper is organized as follows.
Section \ref{sec-related}. reviews background and related work.
Section \ref{sec-infrastructure}. discusses the desired properties of attack-tolerant network
infrastructure.
Section \ref{sec-ft}. describes adversarial fault tolerance on
structured networks.
Section \ref{sec-butterfly}. describes the concurrent multipath routing
algorithm for the butterfly network topology.
Section \ref{sec-discussion}. discusses the our results.
And Section \ref{sec-conclusion}. concludes.

@@

There has been considerable work on trust in network security.
Both centralized and decentralized approaches are commonly used to create
trust infrastructures.
Centralized approaches such as {\em public key infrastructure} (PKI)
suffer from a number of vulnerabilities
\cite{ellison_ten_2000},
which stem largely from the single points of failure inherent to
centralization.
The well-known and widely-used {\em web of trust} approach
\cite{zimmermann_official_1995,ferguson_practical_2003}
is a decentralized alternative.
In a web of trust,
individuals choose who they trust initially.
Trust is then extended to new individuals if they are vouched for by a
currently-trusted individual,
making it possible to quickly establish a large group of trusted nodes.
However, the web of trust approach assumes that trust is infinitely
transitive,
which is unrealistic
\cite{christianson_why_1997}.

Previous work on evaluating incorporating network structure into
trust models has focused on authentication
protocols, showing that independent paths can reduce an adversary's ability
to impersonate a target
\cite{levien_attack-resistant_2009}.
Other work has shown that identifying independent paths in arbitrary networks
is NP-hard and provided approximation algorithms
\cite{reiter_resilient_1998}.
Our work complements these by explicitly quantifying the effects of partial
trust transitivity,
and providing a routing algorithm to efficiently construct independent
paths on a structured network.

Many distributed consensus protocols (such as those used by cryptocurrencies)
are designed to tolerate arbitrary or adversarial faults.
Byzantine agreement protocols
\cite{lamport_byzantine_1982,castro_practical_1999}
provide tolerance against arbitrary faults (including attacks) under
some circumstances, but are limited to small networks due to poor scalability.
Proof-of-work \cite{dwork_pricing_1993,nakamoto_bitcoin:_2008}
and proof-of-stake \cite{king_ppcoin:_2012}
provide better scalability,
but are wasteful of computational and energy resources.
Federated Byzantine Agreement (FBA) \cite{mazieres_stellar_2015}
is scalable, allows for flexible trust,
and is highly fault-tolerant on networks meeting a set of requirements.
However, FBA does not provide a method for evaluating the
fault tolerance properties of different network structures
or for calculating the failure probabilities within a particular network.

There are relatively few attack-tolerance schemes
that focus on network structure,
compared to more popular cryptographic approaches
\cite{ferguson_practical_2003}.
Fiat and Saia described a scheme that combines the butterfly topology
with expander graphs to create a highly censorship-resistant,
content-addressable network \cite{fiat_censorship_2002},
although this scheme exhibits poor scaling due to a very high level of data
replication.
Perhaps the most mature structural solution is the Freenet collaboration
\cite{clarke_freenet:_2001}.
Freenet uses secret sharing
\cite{shamir_how_1979, blakley_safeguarding_1979}
and small-world routing
\cite{zhang_using_2002,kleinberg_small-world_2000}
to create a content-addressable network with a high level of both
confidentiality and censorship resistance.
Freenet guarantees that data is stored redundantly,
but still allows for centralized network structure,
and thus single points of failure,
as data travels from its origin to the redundant storage locations.

@@

Our proposed routing algorithm makes use of a
{\em structured network}, in which link structure is predetermined.
Structured networks have been a popular tool in parallel processing
architectures \cite{kshemkalyani_distributed_2008}.
More recently, peer-to-peer systems based on distributed hash tables have used
structured ``overlay'' networks to map table keys to local TCP/IP routes
\cite{lua_survey_2005,korzun_structured_2013}.
Such networks can be designed to have favorable structural and routing
properties,
which can be used to to guarantee attack-tolerance properties.

@@

\begin{description}
\item[Scalability:]
For large-scale networks, it is important that the infrastructure allows
for the network to grow while remaining functional.
In practice, people, devices, and connections, have limited capabilities
and these limitations need to be considered as part of the design of the
infrastructure.

\item [Decentralization:]
Systems having single points of failure are less tolerant against faults at
those points.
The existence of such points not only increases the likelihood that an attack
will succeed,
but also incentivizes attack by presenting effective targets.
In order to minimize single points of failure,
attack tolerant infrastructures must use {\em both} decentralized protocols and
decentralized network structures.

\item[Stabilizing asymmetry:]
In the context of international conflict,
{\em asymmetric conflicts} allow a
less powerful party to have an advantage over a more powerful party
\cite{mack_why_1975}.
In asymmetric conflicts, the same level of resource expenditure yields different
results for different parties;
the attacker's resources are either more or less effective than the defender's.
We call the latter case stabilizing asymmetry,
because it reduces an attacker's power relative to their target.
With this in mind, an attack-resistant infrastructure will benefit from a high
level of stabilizing asymmetry.

@@

We are specifically concerned with network structure-based approaches to
fault tolerance.
In networks, specific structural properties are required to achieve the
functional properties described above.
\begin{description}
\item [Sparsity and low diameter:]
To achieve scalability, networks must be {\em sparse} and have a
{\em low diameter}.
In practical settings, humans and devices have an upper limit on the number
of connections they can maintain (e.g., Dunbar's number
\cite{dunbar_neocortex_1992}).
In sparse networks, the number of links grows slowly as the network grows in
size, allowing the network to scale without exceeding the nodes' capacity for
links.
Similarly, a low diameter guarantees that as a network grows, a short path will
still exist between any pair of nodes.
While low diameter guarantees a path exists,
paths are only useful if an efficient {\em routing} algorithm exists
to find them.

@@

\item [Redundancy:]
In a network, redundancy refers to the existence of multiple disjoint
paths between nodes or components.
Redundancy can help reduce single points of failure and decrease
centralization.
One measure of redundancy is given by the ratio of edges to nodes
\cite{baran_distributed_1964}.
A single point of failure occurs when a node holds a uniquely
central position.
When alternative paths are added to bypass central nodes,
the network becomes more redundant.

@@

Within the field of {\em fault tolerance},
many techniques have been developed for building reliable systems
out of unreliable components
\cite{avizienis_basic_2004, von_neumann_probabilistic_1956}.
We will make use of standard fault tolerance terminology, summarized here.
A {\em fault} occurs when one component
of a system behaves incorrectly (e.g., a routing node blocks or
alters a message).
The result of that fault (e.g., a recipient receiving conflicting messages)
is an {\em error} state.
If the error is undetected or corrected to the wrong value, the system is
said to have experienced a {\em failure} (e.g., an altered message is
accepted as authentic).
We are concerned in particular with {\em adversarial faults},
which are chosen strategically to maximize the likelihood of a failure.

@@

An important challenge in large-scale, secure communication is the ability to
communicate reliably and securely with remote parties for whom no direct,
trusted link exists.
The commonly-used web of trust approach
\cite{zimmermann_official_1995,ferguson_practical_2003}
extends trust infinitely transitively:
to friends of friends, and friends of friends of friends, and so on.
However, the assumption of infinitely transitive trust is unrealistic
\cite{christianson_why_1997},
and does not allow for the analysis of the effects of network structure.

We can allow for more realistic transitivity assumptions and incorporate
network structure into a trust model with one key insight:
even if no single path between a sender and receiver is fully trusted,
multiple copies of a message can be sent along different paths and compared
by the receiver to detect and correct errors.
However, in the presence of an attacker, there is only benefit in sending an
additional message along a path if that path does not share untrusted
single points of failure with any of the existing message paths
(otherwise, the adversary can compromise both messages
by causing a single fault).
Following \cite{reiter_resilient_1998},
we call paths that do not share any untrusted single points of failure
{\em independent}.
The maximum number of independent paths represents the effective redundancy that
can be utilized by any redundancy-based fault tolerance scheme.

We now propose a method for quantifying the effective redundancy of a network
as a function of its structure and level of trust transitivity.
For the purpose of analyzing adversarial fault tolerance,
this method allows a communication network
to be modeled as a set of virtual channels that directly
connect each node pair,
with each virtual channel providing some level of redundancy.

We now specify the {\em partial trust model}.
We assume an undirected graph $G = (V,E)$,
although the model can easily be extended to directed multigraphs.
Vertices representing commiunicating agents,
and with edges representing mutually trusted communication links.
Let $v \in V$ be an arbitrary sender (Alice)
and $w \in V$ be an arbitrary receiver (Bob).
We assume the presence of an adversary (Mal) who knows the
full structure of the network,
and who can compromise a fixed number of nodes,
gaining complete control of their behavior.
We also assume that Mal is an adversary of Alice and/or Bob specifically,
rather than the network as a whole.
So Mal can compromise any node except for those trusted by Alice or Bob.
We define a {\em trust radius} $h$ such that nodes $u$ and
$u^\prime$ trust each other if their distance is less than $h$.
For a given node $u$,
we call the set of trusted nodes its
{\em trusted neighborhood} $T_h(u)$,
and all nodes at exactly distance $h$ the
{\em trust boundary} $B_h(u)$:
\beq
T_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) < h \right\} \\
B_h(u) &=& \left\{ u^\prime \mid d(u,u^\prime) = h \right\}.
\eeq
The trust boundary $B_h$ plays an important role because these nodes are not
trusted by $u$,
but if compromised can entirely isolate $u$ from the rest of the network.
These trust assumptions imply that when Alice sends a message to Bob,
Mal can only cause faults in the set of nodes outside both of their trusted
neighborhoods: $V \setminus \left(T_h(v) \cup T_h(w)\right)$.
We refer to this set of nodes as the {\em untrusted region}.

Having defined the assumptions of the partial trust model,
we now quantify the {\em effective redundancy} between Alice and Bob
when trust radius $h$ is assumed.
This quantity, $\delta_{v,w,h}$ is exactly the max-flow/min-cut of
the graph after Alice's and Bob's trusted neighborhoods have been
collapsed into single source/sink vertices.
Each trust boundary forms a cut of the network and places an upper bound on the
min-cut:
\beq
\delta_{v,w,h} \leq \min\left( \mid B_h(v) \mid, \mid B_h(w) \mid \right).
\eeq
Equality holds when there are no bottlenecks within the untrusted region,
an indication that the network is decentralized.
The redundancy of the entire graph can be characterized by the minimum over
all vertex pairs:
\beq
\delta_h(V) \equiv \min_{v,w \in V} \delta_{v,w,h}.
\eeq
Thus, for any pair of nodes in the network, at least $\delta_h$ independent,
redundant paths can be constructed between them.
$\delta_h$ is a purely structural network property,
and places an upper bound on the effectiveness of any
redundancy-based fault tolerance scheme.
The more quickly $\delta_h$ grows with $h$,
the better a network is at creating redundancy from trust.
Thus, the scaling of $\delta_h$ can be used to quantify a network's ability
to withstand targeted attacks,
even when the exact trust radius $h$ is unknown.
